Double Standards in Judging Collective Action\thanks{}

### 2022-08-22 Documents/manuscript.tex
update manuscript

\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
      \noindent Collective action is a powerful force driving social
change but often sparks contention about what actions are acceptable
means to effect social change. In four studies (total \emph{N} = 2,246),
we investigated double standards in judging collective action---that is,
whether observers will judge the same protest actions as more acceptable
if the protesters are ingroup members and their cause aligns with the
ingroup's interests (\emph{identity-based double standards}) or if the
protesters' cause aligns with the observers' ideological positions
(\emph{ideology-based double standards}). In two studies, we developed
an instrument of 25 controversial protest actions, based in item
response theory, to measure where people draw the line between
acceptable and unacceptable forms of collective action. In two
preregistered experiments, we used this instrument to test for double
standards in judging collective action for workers' rights in the United
Kingdom and for or against defunding the police in the United States. We
found evidence for ideology-based, but not for identity-based, double
standards: Participants judged the same protest actions to be more
acceptable when the protesters' cause aligned with their own
(system-challenging) ideological position---but showed no ingroup bias
when judging collective action by ingroup and outgroup members. Our
findings have theoretical and practical implications for understanding
the often divided response to prominent social movements.
    \end{abstract}
  \end{@twocolumnfalse}
  % \renewcommand{\abstractname}{Statement of Relevance}
  % \begin{@twocolumnfalse}
  %   \maketitle
  %   \begin{abstract}
  %   \noindent 
  %   \end{abstract}
  % \end{@twocolumnfalse}
]

@@

In this paper, we examine double standards in judging collective
action---that is, whether observers judge the same protest actions as
more or less acceptable depending on their own and the protesters' group
memberships (\emph{identity-based double standards}) or on the
protesters' cause and how it aligns with their own ideological positions
(\emph{ideology-based double standards}). In two studies, we first
develop an instrument of 25 controversial protest actions, based in item
response theory, to capture double standards in judging collective
action. In two preregistered experiments, we then use this instrument to
test for double standards in judging protest actions for workers' rights
in the United Kingdom and for or against defunding the police in the
United States. We find evidence for ideology-based, but not for
identity-based, double standards: People judge the same protest actions
to be more acceptable when the protesters' cause aligns with their own
(system-challenging) ideological positions---but do not show ingroup
bias when judging collective action by ingroup and outgroup members.

@@

Model 1 (M1) estimated varying intercepts for the eight conditions to
test whether, in line with Hypotheses 1a, 1b, or 2a, participants'
responses depended on their own group membership, the protesters' group
membership, or the protesters' cause. Contradicting \emph{Hypothesis
1a}, participants did not consider protest actions performed by their
ingroup to be more acceptable, on average, than the same actions
performed by the relevant outgroup (\(d = 0.01, [-0.04, 0.06]\)).
Contradicting \emph{Hypothesis 1b}, participants did not consider
protest actions for a cause that was nominally aligned with their
ingroup's interests to be more acceptable, on average, than protest
actions for a cause not aligned with their ingroup's interests
(\(d = -0.01, [-0.06, 0.04]\)). Contradicting \emph{Hypothesis 2a},
participants did not consider protest actions for a system-defending
cause to be more acceptable, on average, than protest actions for a
system-challenging cause (\(d = -0.14, [-0.20, -0.09]\)). Instead, we
found that, on average, White participants considered all protest
actions to be more acceptable than Black participants
(\(d = 0.09, [0.04, 0.14]\)) and both Black (\(d = 0.07, [0.03, 0.10]\))
and White (\(d = 0.08, [0.04, 0.11]\)) participants considered the same
actions to be more acceptable when protests were for, rather than
against, defunding the police. As in Experiment 1, we thus found that
participants' responses depended on both the participants' and the
protesters' group memberships---but not in the directions predicted by
our hypotheses.

@@

We explored whether group identification moderated how the participants'
and the protesters' group memberships affected participants' responses.
To that end, we extended Model 1 by estimating participants' responses
as a function of their \emph{z}-standardized identification with their
racial ingroup. As in Model 2, we modeled this relationship with a fixed
and a varying effect. We found, however, that even participants who
strongly identified with their racial ingroup (+1\emph{SD}) did not
consider protest actions performed by their ingroup to be more
acceptable than the same actions performed by the relevant outgroup
(\(d = 0.02, [-0.04, 0.07]\)) and did not consider protest actions for a
cause that was aligned with their ingroup's interests to be more
acceptable than protest actions for a cause not aligned with their
ingroup's interests (\(d = -0.03, [-0.09, 0.02]\)). Our
non-preregistered analyses thus suggested that group identification did
not moderate group differences in judgements about collective action by
ingroup and outgroup members.

@@

@@

Our research was motivated by real-world examples that seemed to suggest
that where people draw the line between acceptable and unacceptable
protest actions depends on \emph{who} the protesters are and \emph{what}
they are protesting---in other words, that there are double standards in
judging collective action. Our findings, however, showed that it is not
so simple. Contradicting our hypothesis of identity-based double
standards, participants in two preregistered experiments did not show
ingroup bias when judging controversial protest actions taken by ingroup
and outgroup members to advance their group's interests. Instead,
supporting our hypothesis of ideology-based double standards,
participants judged the same controversial protest actions to be more
acceptable when the protesters' cause aligned with their own
(system-challenging) ideological position. In the remainder of this
General Discussion, we discuss theoretical and practical implications
for understanding the often divided response to collective action and
consider some limiting conditions on our findings.

Our results showed that, for many controversial protest actions, there
is no consensus on what is and is not an acceptable means to advance a
cause. Instead, this judgement seems to depend on what the cause of a
protest is and how it aligns with one's beliefs and values. In this way,
our results confirm our contention that the distinction between
normative and non-normative forms of collective action should be
situated in the eye of the beholder. By developing an instrument for
measuring where people draw the line between acceptable and unacceptable
forms of collective action, we provided a paradigm that we hope will
stimulate research on when and why people differ in making this
distinction.

After all, this distinction can be consequential. If people believe that
even extreme actions are acceptable means to advance a cause they
support, especially if they moralize their support for that cause
(Mooijman et al., 2018), they might support violent extremism. In this
way, our findings could explain why 45\% of Republicans (2\% of
Democrats) supported the storming of the U.S. Capitol on January 6, 2021
(YouGov, 2021) even though it was illegal, destructive, and violent. If
people believe that even non-violent actions are unacceptable means to
advance a cause they oppose, they might support stifling dissent. In
this way, our findings could explain why Republicans responded to Black
Lives Matter protests by seeking to further criminalize disruptive
protest (e.g., blocking highways, Quinton, 2021) that might be most
effective at challenging injustice (Shuman et al., 2022, 2021). For
those reasons, researchers should further investigate when and why
people apply different standards when judging collective action for
various causes.

@@

Despite its strengths, our research is qualified by several limitations.
First, our research examined the hypothesized relationships for only two
political causes---protecting workers' rights, defunding the
police---and thus provided only limited evidence that our findings
generalize beyond those causes. Relying on few stimuli to establish an
effect threatens the validity and replicability of research findings and
severely constrains psychological research generalize beyond those few
stimuli (Yarkoni, 2022). Future research should sample a wider range of
causes to address this pervasive but often ignored problem (Judd et al.,
2012). This is particularly important when studying collective action as
recent theorizing (Jost et al., 2017) and research (Osborne et al.,
2019) highlighted the importance of differentiating between progressive
and conservative causes. Second, our research was based on two Western,
Educated, Industrialized, Rich, and Democratic (WEIRD; Henrich et al.,
2010) samples, which severely limits the cross-cultural generalizability
of our findings. Future research should include samples from a wider
range of cultural context to examine how the observed relationships vary
across cultural contexts.

@@

@@

@@

@@

>>>>

\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
      \noindent Collective action is a powerful force driving social
change but often sparks contention about what actions are acceptable
means to effect social change. In five studies (total \emph{N} = 2,979),
we investigated double standards in judging collective action---that is,
whether observers will judge the same protest actions as more acceptable
if the protesters are ingroup members and their cause aligns with the
ingroup's interests (\emph{identity-based double standards}) or if the
protesters' cause aligns with the observers' ideological positions
(\emph{ideology-based double standards}). In two studies, we developed
an instrument of 25 controversial protest actions, based in item
response theory, to measure where people draw the line between
acceptable and unacceptable forms of collective action. In three
preregistered experiments, we used this instrument to test for double
standards in judging collective action for workers' rights in the United
Kingdom and for or against defunding the police and for or against
restricting abortion in the United States. We found evidence for
ideology-based, but not for identity-based, double standards:
Participants judged the same protest actions to be more acceptable when
the protesters' cause aligned with their own ideological position---but
showed no ingroup bias when judging collective action by ingroup and
outgroup members. Our findings have theoretical and practical
implications for understanding the often divided response to prominent
social movements.
    \end{abstract}
  \end{@twocolumnfalse}
  % \renewcommand{\abstractname}{Statement of Relevance}
  % \begin{@twocolumnfalse}
  %   \maketitle
  %   \begin{abstract}
  %   \noindent 
  %   \end{abstract}
  % \end{@twocolumnfalse}
]

@@

In this paper, we examine double standards in judging collective
action---that is, whether observers judge the same protest actions as
more or less acceptable depending on their own and the protesters' group
memberships (\emph{identity-based double standards}) or on the
protesters' cause and how it aligns with their own ideological positions
(\emph{ideology-based double standards}). In two studies, we first
develop an instrument of 25 controversial protest actions, based in item
response theory, to capture double standards in judging collective
action. In three preregistered experiments, we then use this instrument
to test for double standards in judging protest actions for workers'
rights in the United Kingdom and for or against defunding the police and
for or against restricting abortion in the United States. We find
evidence for ideology-based, but not for identity-based, double
standards: People judge the same protest actions to be more acceptable
when the protesters' cause aligns with their own ideological
positions---but do not show ingroup bias when judging collective action
by ingroup and outgroup members.

@@

Model 1 (M1) estimated varying intercepts for the eight conditions to
test whether, in line with Hypotheses 1a, 1b, or 2a, participants'
responses depended on their own group membership, the protesters' group
membership, or the protesters' cause. Contradicting \emph{Hypothesis
1a}, participants did not consider protest actions performed by their
ingroup to be more acceptable, on average, than the same actions
performed by the relevant outgroup (\(d = 0.01, [-0.04, 0.06]\)).
Contradicting \emph{Hypothesis 1b}, participants did not consider
protest actions for a cause that was nominally aligned with their
ingroup's interests to be more acceptable, on average, than protest
actions for a cause not aligned with their ingroup's interests
(\(d = -0.01, [-0.06, 0.04]\)). Contradicting \emph{Hypothesis 2a},
participants did not consider protest actions for a system-defending
cause to be more acceptable, on average, than protest actions for a
system-challenging cause (\(d = -0.14, [-0.20, -0.09]\)). Instead, we
found that, on average, White participants considered all protest
actions to be more acceptable than Black participants
(\(d = 0.09, [0.04, 0.14]\)) and both Black (\(d = 0.13, [0.06, 0.20]\))
and White (\(d = 0.16, [0.08, 0.23]\)) participants considered the same
actions to be more acceptable when protests were for, rather than
against, defunding the police. As in Experiment 1, we thus found that
participants' responses depended on both the participants' and the
protesters' group memberships---but not in the directions predicted by
our hypotheses.

@@

We explored whether group identification moderated how the participants'
and the protesters' group memberships affected participants' responses.
To that end, we extended Model 1 by estimating participants' responses
as a function of their \emph{z}-standardized identification with their
racial ingroup. As in Model 2, we modeled this relationship with a fixed
and a varying effect. We found, however, that even participants who
strongly identified with their racial ingroup (+1\emph{SD}) did not
consider protest actions performed by their ingroup to be more
acceptable than the same actions performed by the relevant outgroup
(\(d = 0.01, [-0.00, 0.02]\)) and did not consider protest actions for a
cause that was aligned with their ingroup's interests to be more
acceptable than protest actions for a cause not aligned with their
ingroup's interests (\(d = 0.00, [-0.01, 0.01]\)). Our non-preregistered
analyses thus suggested that group identification did not moderate group
differences in judgements about collective action by ingroup and
outgroup members.

@@

\hypertarget{experiment-3}{%
\section{Experiment 3}\label{experiment-3}}

\hypertarget{method-2}{%
\subsection{Method}\label{method-2}}

We preregistered the sample size, hypotheses, inclusion/exclusion
criteria, measures, and manipulations
(\url{https://osf.io/kphyw/?view_only=59f5792a6d3f4fa48516ea9d5f637822}).
We deviated from the preregistration in two ways. First, we stopped
collecting data before reaching the preregistered sample size. We did so
as the Supreme Court decision on reproductive rights, announced on June
24, 2022, had rendered the experimental manipulation invalid. Second, we
do not report results for two further hypotheses---testing moral
conviction as a potential mechanism underlying ideology-based double
standards---as those findings are beyond the scope of this manuscript
and will be reported in a future publication. We made all materials,
data, and analysis scripts available online
(\url{https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62}).

\hypertarget{study-design-2}{%
\subsubsection{Study Design}\label{study-design-2}}

We used a 2 (quasi-experimental: conservative/liberal participants)
\(\times\) 2 (quasi-experimental: male/female participants) \(\times\) 2
(experimental: for/against restricting abortion) between-subjects design
to test our hypotheses.

\hypertarget{participants-2}{%
\subsubsection{Participants}\label{participants-2}}

We recruited 804 participants from the Prolific subject pool who were 18
years old or older and lived in the United States.\footnote{Data were
  collected between June 10 and 23, 2022. Participants received, on
  average, \$24.70 per hour of participation.} We balanced the sample's
gender and political composition so that half of all participants
identified as conservative and the other half identified as liberal and
so that roughly half of the participants in each group were women and
men. As preregistered, we excluded 71 participants who failed at least
one of five attention checks or reported a different political
orientation than they had reported in the Prolific prescreening
questionnaire. As noted above, we had to stop collecting data before
reaching the preregistered sample size of 800 eligible participants.
This resulted in a final sample of 733 participants
(\(\textit{Mdn} = 37\) years, age range: 18--93 years) of whom 181 were
conservative men, 190 were conservative women, 185 were liberal men, and
177 were liberal women.

\hypertarget{procedure-2}{%
\subsubsection{Procedure}\label{procedure-2}}

After answering where they would you place themselves along the
political spectrum, in line with their response in the Prolific
prescreening questionnaire, participants read the following paragraph:

\begin{quote}
Recent news suggested that the Supreme Court might soon overturn its
\emph{Roe v. Wade} decision that, so far, has guaranteed legal abortions
in the first 20--24 weeks of pregnancy. Overturning \emph{Roe v. Wade}
would allow states to enact laws to further restrict or ban abortions.
\end{quote}

\noindent In the supporting abortion restrictions condition,
participants then read a text presenting arguments for restricting or
banning abortion:

\begin{quote}
Some people welcome this news because they oppose legal abortions. They
argue that life begins at conception and that, therefore, abortion ends
the life of an unborn child. As human life is sacred, ending an unborn
life is wrong and cannot be justified by a pregnant person's right to
choose. In this view, restricting or banning abortions is the only way
to protect unborn lives. We refer to this position as ``supporting
abortion restrictions''.
\end{quote}

\noindent In the opposing abortion restrictions condition, participants
instead read a text presenting arguments against restricting or banning
abortion:

\begin{quote}
Some people are alarmed by this news because they support legal
abortions. They argue that pregnant people have a right to control their
own body and, therefore, have a right to safe and legal abortions.
Restricting or banning abortion would lead to more illegal and unsafe
abortions. In this view, providing legal abortions is the only way to
protect pregnant people's rights and to prevent harm. We refer to this
position as ``opposing abortion restrictions''.
\end{quote}

\noindent Participants were then asked to select which position the text
described and what arguments for that position were presented in the
text. If they selected the wrong answer, they were instructed to reread
the text and select the right answer. Participants then rated to what
extent they support or oppose restricting abortion and to what extent
their feelings about this issue are a moral conviction.

Participants then read about protesters protesting either for (in the
supporting abortion restrictions condition) or against (in the opposing
abortion restrictions condition) restricting or banning abortion:

\begin{quote}
Earlier, you read why some people {[}support/oppose{]} further
restricting or banning abortion. Some local residents who hold this view
want to protest {[}for/against{]} restricting or banning abortion.
\end{quote}

\noindent Participants rated which of several protest actions they
consider acceptable means to protest for or against restricting
abortion. On the next pages, participants completed the remaining
measures. On the final page, participants responded to two attention
checks that assessed whether they paid attention to the experimental
manipulation: ``In this study, you read about some people's view on a
policy change. What was the policy change?'' and ``In this study, you
then read about some protesters. Were the protesters for or against
restricting abortion?''.

\hypertarget{measures-2}{%
\subsubsection{Measures}\label{measures-2}}

We measured the outcome variable by asking participants to think about
the protesters they had read about and to decide, for each of 25 protest
actions presented in a randomized order, whether they thought this
action was ``an acceptable means for them to protest {[}for/against{]}
restricting or banning abortion'' (1 = \emph{yes}, 0 = \emph{no}; see
Table C1).

We included additional measures to describe the sample, to report as
descriptive statistics, to use in non-preregistered analyses or future
analyses not reported in this paper.

We measured support for restricting abortion with one item: ``Do you
support or oppose restricting (or banning) abortion?'' (1 =
\emph{strongly oppose}, 5 = \emph{strongly support}). We also included a
three-item measure of moral conviction (Ryan, 2014) that assessed to
what extent participants' feelings about restricting or banning abortion
were, for example, connected to their core moral beliefs or convictions
(1 = \emph{not at all}, 5 = \emph{very much}).

We measured gender-related system-justifying beliefs with eight items
(adapted from Jost \& Kay, 2005), for example, ``in general, relations
between men and women are fair'' (1 = \emph{strongly disagree}, 7 =
\emph{strongly agree}; McDonald's \(\omega = .93\)). A confirmatory
factor analysis model in which all items loaded onto a single factor
showed acceptable fit,
\(\chi^2 (20) = 111.90; \text{CFI} = 0.97; \text{TLI} = 0.95; \text{RMSEA} = 0.09, [0.08, 0.11]\).

We will measured moral concerns with the 36-item moral foundations
questionnaire (MFQ-2, Atari et al., 2022) which assesses to what extent
participants endorse concerns about care, equality, proportionality,
loyalty, authority, and purity (e.g., ``I believe chastity is an
important virtue.''; 1 = \emph{does not describe me at all}, 5 =
\emph{describes me extremely well}). We embedded three further attention
checks within the questionnaire (e.g., ``To show that you are paying
attention and giving your best effort, please select `moderately
describes me'.'').

In addition to demographic questions, we asked participants to what
extent they identified with their gender (1 = \emph{strongly disagree},
7 = \emph{strongly agree}) and measured political orientation with a
one-item semantic differential scale: ``People often describe their
political orientation as liberal or conservative. On a scale from
liberal to conservative, where would you position yourself?'' (1 =
\emph{liberal}, 7 = \emph{conservative}).

\hypertarget{results-2}{%
\subsection{Results}\label{results-2}}

\hypertarget{reactions-to-the-manipulation-2}{%
\subsubsection{Reactions to the
Manipulation}\label{reactions-to-the-manipulation-2}}

As expected, participants who identified as conservative reported being
far more conservative (\(M = 6.05, SD = 0.79\)) than participants who
identified as liberal
(\(M = 1.73, SD = 0.80; \text{Cohen's}~d = 5.43\)). Likewise,
conservative participants tended to endorse system-justifying beliefs
(\(M = 5.10, SD = 0.95\)) while liberal participants tended to reject
system-justifying beliefs (\(M = 2.86, SD = 1.03; d = 2.26\)). As
expected, conservative and liberal participants differed in their
reactions to the manipulation as the former tended to support
restricting abortion (\(M = 3.77, SD = 1.48\)) while the latter tended
to oppose restricting abortion (\(M = 1.28, SD = 0.85; d = 2.06\)). When
comparing politically balanced subsamples of men and women, women
(\(M = 2.55, SD = 1.76\)) did not differ from men
(\(M = 2.55, SD = 1.72\)) in their attitudes toward restricting abortion
(\(d = -0.02\)).

\hypertarget{preregistered-analyses-2}{%
\subsubsection{Preregistered Analyses}\label{preregistered-analyses-2}}

\begin{figure*}[!t]
\centering
\caption{Results from the preregistered analyses for Experiment 3}
\includegraphics[scale=1]{../Experiment 3/figures/figure-6}
\caption*{\textit{Note.} Results show the estimated \textit{z}-standardized tendency ($\theta_{(z)}$) to consider more controversial actions acceptable means to protest for or against restricting abortion as a function of the participants' ideology and gender.}
\label{fig:f6}
\end{figure*}

To test our hypotheses, we used a two-parameter logistic item response
theory model, identical to the one used in Experiment 2, to estimate how
likely participants were to consider each action an acceptable means of
protest as a function of the protesters' cause and of the participants'
gender and ideology. Figure \ref{fig:f6} shows the results of our
preregistered analyses.

As in Experiments 1 and 2, we found no evidence for identity-based
double standards: Female participants did not consider protest actions
\emph{against} restricting abortion to be any more acceptable, on
average, than male participants or protest actions \emph{for}
restricting abortion (\(d = 0.14, [0.00, 0.29]\)). Instead, we again
found evidence for ideology-based double standards: Conservative
participants considered the same protest actions to be more acceptable,
on average, when protesters \emph{supported} restricting abortion
(\(d = 0.24, [0.07, 0.43]\)) while liberal participants considered the
same protest actions to be more acceptable when protesters
\emph{opposed} restricting abortion (\(d = 0.90, [0.72, 1.08]\)). This
pattern was consistent across subsamples as conservative women
(\(d = 0.30, [0.04, 0.55]\)) and, to a lesser extent, conservative men
(\(d = 0.19, [-0.07, 0.46]\)) considered the same protest actions to be
more acceptable when protesters \emph{supported} restricting abortion
and as both liberal men (\(d = 0.76, [0.49, 1.03]\)) and women
(\(d = 1.04, [0.78, 1.32]\)) considered the same protest actions to be
more acceptable when protesters \emph{opposed} restricting abortion.
This double standard was, however, more pronounced among liberal
participants (\(d = 0.66, [0.41, 0.92]\)). In this way, Experiment 3
provided evidence for both ideological symmetry and asymmetry in judging
collective action for and against restricting abortion.

\hypertarget{non-preregistered-analyses-2}{%
\subsubsection{Non-Preregistered
Analyses}\label{non-preregistered-analyses-2}}

\hypertarget{discussion-2}{%
\subsection{Discussion}\label{discussion-2}}

@@

Our research was motivated by real-world examples that seemed to suggest
that where people draw the line between acceptable and unacceptable
protest actions depends on \emph{who} the protesters are and \emph{what}
they are protesting---in other words, that there are double standards in
judging collective action. Our findings, however, showed that it is not
so simple. Contradicting our hypothesis of identity-based double
standards, participants in two preregistered experiments did not show
ingroup bias when judging controversial protest actions taken by ingroup
and outgroup members to advance their group's interests. Instead,
supporting our hypothesis of ideology-based double standards,
participants judged the same controversial protest actions to be more
acceptable when the protesters' cause aligned with their own
(system-challenging) ideological position. In the remainder of this
General Discussion, we consider some limiting conditions on our findings
and discuss theoretical and practical implications for understanding the
often divided response to collective action.

@@

Despite its strengths, our research is qualified by important
limitations. First, our research examined the hypothesized relationships
for three political causes---protecting workers' rights, defunding the
police, restricting abortion---and thus provided only limited evidence
that our findings generalize beyond those causes. Relying on few stimuli
to establish an effect threatens the validity and replicability of
research findings and constrains the generalizability of psychological
research (Yarkoni, 2022). Future research should sample a wider range of
causes to address this pervasive but often ignored problem (Judd et al.,
2012). This is particularly important when studying collective action as
recent theorizing (Jost et al., 2017) and research (Osborne et al.,
2019) highlighted the importance of differentiating between progressive
and conservative causes. Second, our research was based on samples from
two Western, Educated, Industrialized, Rich, and Democratic (WEIRD;
Henrich et al., 2010) contexts---the United Kingdom and the United
States---which limits the cross-cultural generalizability of our
findings.

Most broadly, our results showed that, for many controversial protest
actions, there is no consensus on what is and is not an acceptable means
to advance a cause. Instead, this judgement seems to depend on what the
cause of a protest is and how it aligns with one's beliefs and values.
In this way, our results confirm our contention that the distinction
between normative and non-normative forms of collective action should be
situated in the eye of the beholder. By developing an instrument for
measuring where people draw the line between acceptable and unacceptable
forms of collective action, we provided a paradigm that we hope will
stimulate research on when and why people differ in making this
distinction.

After all, this distinction can be consequential. If people believe that
even extreme actions are acceptable means to advance a cause they
support, especially if they moralize their support for that cause
(Mooijman et al., 2018), they might support violent extremism. In this
way, our findings could explain why 45\% of Republicans supported the
storming of the U.S. Capitol on January 6, 2021 (YouGov, 2021) even
though it was illegal, destructive, and violent. If people believe that
even non-violent actions are unacceptable means to advance a cause they
oppose, they might support stifling dissent. In this way, our findings
could explain why Republicans responded to Black Lives Matter protests
by seeking to further criminalize disruptive protest (e.g., blocking
highways, Quinton, 2021) that might be most effective at challenging
injustice (Shuman et al., 2022, 2021). For those reasons, we encourage
researchers to use our paradigm to further investigate when and why
people apply different standards when judging collective action for
different causes.

@@

\leavevmode\vadjust pre{\hypertarget{ref-atari_morality_2022}{}}%
Atari, M., Haidt, J., Graham, J., Koleva, S., Stevens, S. T., \&
Dehghani, M. (2022). Morality beyond the {WEIRD}: How the nomological
network of morality varies across cultures. \emph{PsyArXiv}.
\url{https://doi.org/10.31234/osf.io/q6c9r}

@@

\leavevmode\vadjust pre{\hypertarget{ref-jost_exposure_2005}{}}%
Jost, J. T., \& Kay, A. C. (2005). Exposure to benevolent sexism and
complementary gender stereotypes: Consequences for specific and diffuse
forms of system justification. \emph{Journal of Personality and Social
Psychology}, \emph{88}(3), 498--509.
\url{https://doi.org/10.1037/0022-3514.88.3.498}

@@

\leavevmode\vadjust pre{\hypertarget{ref-ryan_reconsidering_2014}{}}%
Ryan, T. J. (2014). Reconsidering moral issues in politics. \emph{The
Journal of Politics}, \emph{76}(2), 380--397.
\url{https://doi.org/10.1017/S0022381613001357}

@@

\setcounter{table}{0}
\renewcommand{\thetable}{C\arabic{table}}

\section{Appendix C}

\caption{List of protest actions used in Experiment 3}
\small
\begin{tabularx}{\linewidth}{rXr}
\addlinespace
\toprule
\# & Action & Pr\\
\midrule
1 & make a public speech & 94\%\\
2 & hand out flyers, leaflets, or pamphlets & 92\%\\
3 & wear or display political symbols (e.g., patches, flags, bumper stickers) & 92\%\\
4 & hold meetings to influence the public & 92\%\\
5 & attend or organise a protest march & 91\%\\
6 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 91\%\\
7 & donate to activist groups & 90\%\\
8 & refuse to buy goods or services from companies that advocate [against/for] restricting or banning abortion (boycott) & 90\%\\
9 & paste up posters with political messages in places where it is allowed and encouraged & 90\%\\
10 & join or form a group of activists & 90\%\\
11 & donate to politicians who advocate [for/against] restricting or banning abortion & 89\%\\
12 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 84\%\\
13 & refuse to accept honors or awards in protest & 82\%\\
14 & refuse to work (strike) & 56\%\\
15 & visit people in their homes to convince them about the issue (canvassing, door knocking) & 50\%\\
16 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 41\%\\
17 & attend a protest march even though it might be unlawful & 40\%\\
18 & \textit{protest outside the homes of politicians who advocate [against/for] restricting or banning abortion} & 40\%\\
19 & attend a protest march even though it might turn violent & 40\%\\
20 & paste up posters with political messages in places where it is not allowed or encouraged & 26\%\\
21 & enter and refuse to leave a building (occupation) & 25\%\\
22 & refuse to pay fees, fines, or taxes in protest & 23\%\\
23 & mock or insult individuals who are [against/for] restricting or banning abortion & 20\%\\
24 & spray paint political messages in public places & 15\%\\
25 & disrupt traffic (e.g., blocking roads) & 12\%\\
\bottomrule
\end{tabularx}
\caption*{\textit{Note.} Actions are ordered by the proportion of participants, across all conditions, who considered it to be an acceptable means of protest in Experiment 3. Action in \textit{italics} replaced an action used in Experiment 2 because it better fit the context of Experiment 3.}
